---
title: "542 Deliverable 2"
date: "2.11.23"
output:
  html_document:
    df_print: paged
---
Loading in data and identifying the data types
```{r, message=FALSE}

location ='https://github.com/Drops-of-Jupyter/Deliverable-2/raw/main/'
file ='finaldata.csv'
datalink = paste0(location,file)

deliv2data = read.csv(datalink)

str(deliv2data)
```
Setting the random seed for replication purposes
```{r, message=FALSE}
set.seed(999)
```
Adjusting the data to be better fit for analysis
```{r, message=FALSE}
deliv2data[!complete.cases(deliv2data),-c(1,2)]
```

```{r}
deliv2data=deliv2data[complete.cases(deliv2data),]
```
Looking at value ranges
```{r}
boxplot(deliv2data[,-c(1,2)])
```
Adjusting scale 
```{r}
deliv2data[,-c(1,2)]=as.data.frame(scale(deliv2data[,-c(1,2)]))
```
Replotting box plot to see value ranges
```{r}
boxplot(deliv2data[,-c(1,2)])
```
Subsetting the data and checking it worked
```{r, message=FALSE}
dataToCluster=deliv2data[,-c(1,2)]
row.names(dataToCluster)=deliv2data$country
```
Deciding distance method and computing the distance matrix (using gower measurement)
```{r, message=FALSE}
library(cluster)
distanceMatrix=daisy(x=dataToCluster, metric = "gower")
```
Producing a 2 dimensional map of the data
```{r, message=FALSE}
projectedData = cmdscale(distanceMatrix, k=2)
```

```{r, message=FALSE}
##saving coordinates to original dataframe
deliv2data$dim1 = projectedData[,1]
deliv2data$dim2 = projectedData[,2]
##looking at some of the coordinates
deliv2data[,c('dim1','dim2')][1:10,]
```

```{r}
##Creating the map
library(ggplot2)
base= ggplot(data=deliv2data,
             aes(x=dim1, y=dim2,
                 label=country)) 
base + geom_text(size=2)
```
Producing a dendogram as an alternate visualization
```{r}
hc = hclust(distanceMatrix)
plot(hc,hang = -1,cex=0.5)
```

Computing clustering suggestions to see how many clusters to make
```{r, message=FALSE}
##Agglomerative
library(factoextra)
```


```{r}
fviz_nbclust(dataToCluster, 
             hcut,
             diss=distanceMatrix,
             method = "gap_stat",
             k.max = 10,
             verbose = F,
             hc_func = "agnes")
##divisive
fviz_nbclust(dataToCluster, 
             hcut,
             diss=distanceMatrix,
             method = "gap_stat",
             k.max = 10,
             verbose = F,
             hc_func = "diana")

```
Computing the clusters for both meaasurement types
```{r}

NumberOfClusterDesired=3

res.agnes= hcut(distanceMatrix, 
                k = NumberOfClusterDesired,
                isdiss=TRUE,
                hc_func='agnes',
                hc_method = "ward.D2")

# Hierarchical technique- divisive approach
res.diana= hcut(distanceMatrix, 
                k = NumberOfClusterDesired,
                isdiss=TRUE,
                hc_func='diana',
                hc_method = "ward.D2")
```
Saving results to data frame
```{r}
deliv2data$agn=as.factor(res.agnes$cluster)
deliv2data$dia=as.factor(res.diana$cluster)
```
Verifying ordinality
```{r}
aggregate(data=deliv2data,
          death.rate~agn,
          FUN=mean)
```

```{r}
aggregate(data=deliv2data,
          death.rate~dia,
          FUN=mean)
```
reordering to be smallest to greatest 
```{r, message=FALSE}
library(dplyr)

deliv2data$dia=dplyr::recode_factor(deliv2data$dia, 
                  `1`='1',`2`='3',`3`='2',.ordered = T)
```
creating silhouettes to see what measurement is best
```{r, message=FALSE}
fviz_silhouette(res.agnes)

##silhouette width =.42
```

```{r, message=FALSE}
fviz_silhouette(res.diana)

##Silhouette width = .5
```
checking for how many countries are poor in each measurement 
```{r}
agnEval=data.frame(res.agnes$silinfo$widths)
diaEval=data.frame(res.diana$silinfo$widths)

agnPoor=rownames(agnEval[agnEval$sil_width<0,])
diaPoor=rownames(diaEval[diaEval$sil_width<0,])

```

```{r}
diaPoor

#3 countries
```


```{r}
agnPoor

#3 countries
```
comparing visually
```{r}

#Agnes
base= ggplot(data=deliv2data,
             aes(x=dim1, y=dim2,
                 label=country)) 
agnPlot=base + labs(title = "AGNES") + geom_point(size=2,
                                              aes(color=agn),
                                              show.legend = T) 
#Diana
diaPlot=base + labs(title = "DIANA") + geom_point(size=2,
                                              aes(color=dia),
                                              show.legend = T) 

library(ggpubr)

ggarrange(agnPlot, diaPlot,ncol = 2,common.legend = T)

```
adding labels to the visual
```{r}
LABELdia=ifelse(deliv2data$country%in%diaPoor,deliv2data$country,"")
LABELagn=ifelse(deliv2data$country%in%agnPoor,deliv2data$country,"")
```


```{r}
library(ggrepel)
diaPlot=diaPlot + 
        geom_text_repel(aes(label=LABELdia),
                        max.overlaps=50,
                        min.segment.length =unit(0,'lines'))
```

```{r}
agnPlot= agnPlot + 
         geom_text_repel(aes(label=LABELagn),
                         max.overlaps = 50,
                         min.segment.length = unit(0, 'lines'))
```

```{r}
ggarrange(agnPlot, 
          diaPlot,
          ncol = 2,
          common.legend = T)
```

Creating dendograms of both agnes and diana measurements
```{r, message=FALSE}
fviz_dend(res.agnes,
          k=NumberOfClusterDesired, 
          cex = 0.45, 
          horiz = T,
          main = "AGNES approach")
```

```{r}
fviz_dend(res.diana,
          k=NumberOfClusterDesired, 
          cex = 0.45, 
          horiz = T,
          main = "DIANA approach")
```


Both measurement types are viable options. I believe based on the results of the visual tests conducted and the shiloutte width of .5, that Diana would be the best fitting measurement type. 
